---
title: "Multiple Patient Prediction"
output: html_document
---


```{r include=FALSE}
source('./utils/utils.R')
source('./utils/multiple_patient_prediction_utils.R')
source('./model/israeli_multi_state_competing_risks_model.R')
library(ggplot2)
library(formattable)

set.seed(42)

```

# Introduction

This notebook provides a running example for predicting the following for a set of currently hospitalized patients:

1. The number of hospitalized patients per day
2. The number of patients in the Severe state (named "Critical" in the paper) 
3. The number of deceased patients per 5 day block

We also show how you could test different future scenarios by providing a set of future hospitalizations. 


# Initializations:

### Load pre-fit model:

```{r}
load('./cache/multiple_patient_prediction_model.Rdata')
```

### Set Monte Carlo Parameters

```{r}
# the number of simulated runs per patient, 10K is our recommendation
N_MONTE_CARLO_RUNS = 100#*1000

# the maximum number of transitions in the Multi-State model
# this is used to exclude outlier paths with more transitions than
# the maximal number of transitions observed in the data
MAX_PATH_LENGTH = 9 
```


### Define `T`: the day from which we predict

```{r}
T = Sys.Date() # use the date of today
PREDICT_N_DAYS_AHEAD = 30 # predict one month ahead
```


# Load Hospitalized Patient Data:

In this example we generate random data, but you can use your own data by reading a dataframe which complies with the following conventions:

1. `sex`: MALE (1), FEMALE (0)
2. `age`: in years
3. `state_at_hospitalization`: either MILD (2), MODERATE (3) or SEVERE (4). Note, these correspond with the states MODERATE, SEVERE, CRITICAL which are used in the paper
4. `date_of_hospitalization`: date in the format 'yyyy-mm-dd'
5. `current_state`: same values as 3 above
6. `days_spent_in_current_state`: non-negative integer
7. `was_severe`: Binary indicator. Visited severe state at least once before entering the current state (1), or not (0)


```{r}
hospitalized_patients_df = generate_random_hospitalized_patients_df(n_hospitalized_patients = 30,
                                                                    T = T,
                                                                    PREDICT_N_DAYS_AHEAD = PREDICT_N_DAYS_AHEAD)

# alternatively:
# hospitalized_patients_df = read.csv('PATH TO YOUR PATIENT DATA CSV')

formattable(head(hospitalized_patients_df, n=5))
```

### Convert Table into the Format Acceptable by the Model:

```{r}
hospitalized_patients = lapply(1:nrow(hospitalized_patients_df), function(row_idx) {
  
  row = hospitalized_patients_df[row_idx,]
  
  construct_patient_object(
    sex = row$sex, 
    age = row$age, 
    state_at_hospitalization = row$state_at_hospitalization,
    date_of_hospitalization = row$date_of_hospitalization,
    current_state = row$current_state,
    days_spent_in_current_state = row$days_spent_in_current_state,
    was_severe  = row$was_severe,
    T=T
  )
})
```


## Generate a Random Set of Future Arrivals:

```{r}
future_arrival_patients_df = generate_future_arrival_patients_df(n_future_arrivals = 100,
                                                                 T = T,
                                                                 PREDICT_N_DAYS_AHEAD = PREDICT_N_DAYS_AHEAD)

# alternatively:
# future_arrival_patients_df = read.csv('PATH TO YOUR FUTURE ARRIVAL SCENARIO CSV')

formattable(head(future_arrival_patients_df, n=5))
```

```{r}
future_arrival_patients = lapply(1:nrow(future_arrival_patients_df), function(row_idx) {
  
  row = future_arrival_patients_df[row_idx, ]
  
  construct_future_arrival_patient_object(
    sex = row$sex,
    age = row$age,
    state_at_hospitalization = row$state_at_hospitalization,
    date_of_hospitalization = row$date_of_hospitalization
  )  
})
```


### Merge the two lists of patients, and perform a monte carlo simulation per patient

```{r}
INCLUDE_FUTURE_ARRIVAL_PATIENTS = TRUE
```


```{r}
all_patients = hospitalized_patients
if (INCLUDE_FUTURE_ARRIVAL_PATIENTS) all_patients = append(all_patients, 
                                                           future_arrival_patients)  


all_patients = run_monte_carlo_per_patient(all_patients, 
                                           N_MONTE_CARLO_RUNS, 
                                           MAX_PATH_LENGTH)
```

# Plot the number of hospitalized patients per day 

The plot below displays the mean number of hospitalized patients, over all monte carlo simulations. 

For each day, the range between the lower and upper quantiles is filled in red. In the example below we display the 25 and 75 percentiles. 

```{r echo=FALSE}
plot_expected_number_of_patients_each_day(all_patients,
                                          first_date = T,
                                          last_date = T + PREDICT_N_DAYS_AHEAD,
                                          states_to_include = c(SEVERE, MILD_OR_MODERATE), # all "hospitalized" states
                                          lower_quantile = 0.25,
                                          upper_quantile = 0.75)

```


# Plot the number of patients hospitalized at SEVERE state per day 

The plot below displays the mean number of hospitalized patients **in the SEVERE state** (named Critical in the paper), over all monte carlo simulations. 

For each day, the range between the lower and upper quantiles is filled in red. In the example below we display the 25 and 75 percentiles.

```{r}
plot_expected_number_of_patients_each_day(all_patients,
                                          first_date = T,
                                          last_date = T + PREDICT_N_DAYS_AHEAD,
                                          states_to_include = c(SEVERE),
                                          lower_quantile = 0.25,
                                          upper_quantile = 0.75)

```



# Expected Number of Deaths


```{r}
formattable(construct_expected_deaths_table(all_patients, 
                                            T, 
                                            PREDICT_N_DAYS_AHEAD))
```


# 