---
title: "Multiple Patient Prediction"
output: html_document
---


```{r include=FALSE}
source('./utils/utils.R')
source('./utils/multiple_patient_prediction_utils.R')
source('./model/israeli_multi_state_competing_risks_model.R')
library(ggplot2)
library(formattable)

set.seed(42)

```

# Introduction

This notebook provides a running example for predicting the following for a set of currently hospitalized patients:

1. The expected number of hospitalized patients per day.
2. The expected number of patients in a Critical state (named "Critical" in the paper, but `SEVERE` in code).
3. The expected number of deceased patients by days since hospitalization, in 5 day intervals.

We also show how you could test different future scenarios by loading a set of future hospitalizations you defined. 


# Initializations:

### Load pre-fit model:

```{r}
load('./cache/multiple_patient_prediction_model.Rdata')
```

### Set Monte Carlo Parameters

```{r}
# the number of simulated runs per patient, 10K is our recommendation
N_MONTE_CARLO_RUNS = 10*1000

# the maximum number of transitions in the Multi-State model
# this is used to exclude outlier paths with more transitions than
# the maximal number of transitions observed in the data
MAX_PATH_LENGTH = 9 
```


### Define `T`: the day from which we predict

```{r}
T = Sys.Date() # use the date of today
PREDICT_N_DAYS_AHEAD = 30 # predict one month ahead
```


# Load Hospitalized Patient Data:

In this example we generate **random stub data**, but you can use your own data by reading a dataframe which complies with the following conventions:

1. `sex`: MALE (1), FEMALE (0)
2. `age`: in years
3. `state_at_hospitalization`: either `MILD` (2), `MODERATE` (3) or `SEVERE` (4). Note, these correspond with the states Moderate, Severe, Critical which are used in the paper
4. `date_of_hospitalization`: date in the format 'yyyy-mm-dd'
5. `current_state`: same values as 3 above
6. `days_spent_in_current_state`: non-negative integer
7. `was_severe`: Binary indicator. Visited severe state at least once before entering the current state (1), or not (0)

**Note:** The data generated here is not intended to reflect any true distribution of patients but only serves as an example of the needed data **format**. Do not base predictions on this stub data.


```{r}
hospitalized_patients_df = generate_random_hospitalized_patients_df(n_hospitalized_patients = 30,
                                                                    T = T,
                                                                    PREDICT_N_DAYS_AHEAD = PREDICT_N_DAYS_AHEAD)

# alternatively:
# hospitalized_patients_df = read.csv('PATH TO YOUR PATIENT DATA CSV')

formattable(head(hospitalized_patients_df, n=5))
```

### Convert Table into the Format Acceptable by the Model:

```{r}
hospitalized_patients = lapply(1:nrow(hospitalized_patients_df), function(row_idx) {
  
  row = hospitalized_patients_df[row_idx,]
  
  construct_patient_object(
    sex = row$sex, 
    age = row$age, 
    state_at_hospitalization = row$state_at_hospitalization,
    date_of_hospitalization = row$date_of_hospitalization,
    current_state = row$current_state,
    days_spent_in_current_state = row$days_spent_in_current_state,
    was_severe  = row$was_severe,
    T=T
  )
})
```


## Load Future Arrival Data:

**Note:** The stub data generated here is not intended to reflect any true distribution of patients but only serves as an example of the needed data **format**. Do not base predictions on this stub data.

```{r}
future_arrival_patients_df = generate_future_arrival_patients_df(n_future_arrivals = 100,
                                                                 T = T,
                                                                 PREDICT_N_DAYS_AHEAD = PREDICT_N_DAYS_AHEAD)

# alternatively:
# future_arrival_patients_df = read.csv('PATH TO YOUR FUTURE ARRIVAL SCENARIO CSV')

formattable(head(future_arrival_patients_df, n=5))
```

```{r}
future_arrival_patients = lapply(1:nrow(future_arrival_patients_df), function(row_idx) {
  
  row = future_arrival_patients_df[row_idx, ]
  
  construct_future_arrival_patient_object(
    sex = row$sex,
    age = row$age,
    state_at_hospitalization = row$state_at_hospitalization,
    date_of_hospitalization = row$date_of_hospitalization
  )  
})
```


### Merge the two lists of patients, and perform a monte carlo simulation per patient

```{r}
INCLUDE_FUTURE_ARRIVAL_PATIENTS = TRUE
```


```{r}
all_patients = hospitalized_patients
if (INCLUDE_FUTURE_ARRIVAL_PATIENTS) all_patients = append(all_patients, 
                                                           future_arrival_patients)


all_patients = run_monte_carlo_per_patient(all_patients, 
                                           N_MONTE_CARLO_RUNS, 
                                           MAX_PATH_LENGTH)
```

# Plot the number of hospitalized patients per day 

The plots below displays the mean number of hospitalized patients, over all monte carlo simulations.

For each day, we plot:

* mean number of hospitalized patients
* 25 to 75 quantiles of number of hospitalized patients (red fill)
* 10 to 90 quantiles of number of hospitalized patients (orange fill)


```{r echo=FALSE}
plot_expected_number_of_patients_each_day(all_patients,
                                          first_date = T,
                                          last_date = T + PREDICT_N_DAYS_AHEAD,
                                          states_to_include = c(SEVERE, MILD_OR_MODERATE)) # all "hospitalized" states
```


# Plot the number of patients hospitalized at SEVERE state per day 

The plot below is simialr to that above, except it displays the mean number of hospitalized patients **in a Critical state** (`SEVERE`)

```{r}
plot_expected_number_of_patients_each_day(all_patients,
                                          first_date = T,
                                          last_date = T + PREDICT_N_DAYS_AHEAD,
                                          states_to_include = c(SEVERE))
```



# Expected Number of Deaths

Below you can see the expected proportion of deaths by days since hospitalization. That is, we are counting deaths among all patients 5 days since hospitalization, 10 days hospitalization etc.; averaging over all sampled monte carlo paths.

This analysis includes future arrival patients if `INCLUDE_FUTURE_ARRIVAL_PATIENTS = TRUE`

Note: deaths occuring after `T + PREDICT_N_DAYS_AHEAD` are excluded from the displayed statistics. 

```{r}
formattable(expected_deaths_by_days_since_hospitalization(all_patients, 
                                                          T,
                                                          PREDICT_N_DAYS_AHEAD))
```
