---
title: "Single Patient Prediction"
output: html_document
---


```{r include=FALSE}
library(ggplot2)

source('./utils/utils.R')
source('./utils/single_patient_prediction_utils.R')
library(formattable)


# load pre-fit models:
load('./cache/single_patient_prediction_environment_1.RData')
load('./cache/single_patient_prediction_environment_2.RData')
weighted_bootstrap_models = c(weighted_bootstrap_models_first_50, weighted_bootstrap_models_last_50)
remove(weighted_bootstrap_models_first_50, weighted_bootstrap_models_last_50)
```


# Introduction

This notebook provides a running example of estimating the following values, for a single patient: 

  * probability of death
  * probability of future visit to critical state
  * quantiles of predicted time in hospital
  * quantiles of predicted time in critical state
  * Cumulative Distribution Function (CDF) for remaining time at hospital

**Note:** In the code, the constants denoting the states are named `MILD`, `MODERATE` and `SEVERE`. In the paper these constants correspond with the states named Moderate, Severe and Critical, respectively.

# Follow These Steps:

## 1. Define Patient Covariates:

We shall provide estimates for a 55 year old male hospitalized in a critical condition. Predictions below are done from time of hospitalization but estimates can be predicted from any time since hospitalization.


### Sex

Either `MALE` or `FEMALE`

```{r}
sex = MALE
```

### Age (in years)

```{r}
age = 55
```


### Patient's State at Hospitalization

Denoted as `entry_state`, this can be either MILD, MODERATE or SEVERE.

NOTE: in the paper, these are named MODERATE, SEVERE and CRITICAL respectively.

```{r}
entry_state = SEVERE
```


### Patient's Current State

For prediction from time of hospitalization this is necessarily the same as entry_state, but if we predict from some time during hospitalization this can be another state.

```{r}
current_state = SEVERE
```


### Days since hospitalization until entry to current state

Integer number of days since hospitalization until the patient entered the current state

Examples:

 1. For prediction from day of hospitalization (necessarily from first state) use: 0
 
 2. If the patient was hospitalized for 10 days, 3 of them spent in the current state, then 7 days have passed since hospitalization until entry to the curent state, use: 7

```{r}
days_since_hospitalization_until_entry_to_current_state = 0
```

### Days spent at current state

Integer number of days spent at the current state

Examples:

  1. For prediction from day of hospitalization (necessarily from first state) use: 0
  
  2. If patient was hospitalized for 10 days, and is already 3 days in the current state, use: 3

```{r}
days_at_current_state = 0
```

From the two definitions above we have:

```{r}
total_days_since_hospitalization = days_since_hospitalization_until_entry_to_current_state + 
                                   days_at_current_state
```


### Indicator Variable for Previous Visit to `SEVERE` State:

A binary 0/1 indicator of a **past** visit to the `SEVERE` state (In the paper: "Critical"). 1 indicates there was such a visit. 

Note: if the patient was hospitalized in `SEVERE` state, the indicator should initially be 0, as there was no **past** visit to the `SEVERE` state.

```{r}
was_severe = 0
```



### Construct The Covariates:

```{r}
covariates = construct_covariates(sex = sex,
                                  age = age,
                                  entry_state = entry_state,
                                  cumulative_time = days_since_hospitalization_until_entry_to_current_state,
                                  was_severe = was_severe)
```



## 2. Decide if you wish to compute confidence intervals 

This notebook computes 95% confidence intervals for the estimates listed above. Intervals are estimated using a weighted bootstrap based on 100 models, i.e 100 models each running 20K monte carlo simulations. This can be relatively slow.

Note: For estimates of the different quantiles of hospitalization time remaining and time remaining in a critical state, in estimates of extreme quantiles, coverage could possibly differ from 95%.

```{r}
COMPUTE_CONFIDENCE_INTERVALS = FALSE
```


## 3. Set Monte Carlo Paramaters:


### Define the number of monte carlo paths to sample:

```{r}
M_MONTE_CARLO_PATHS = 20*1000 # recommended
```


### Set maximum state transitions

This is used to exclude outlier paths. For instance, this rule can be used to exclude occurences of paths longer than those observed in the data.

```{r}
MAX_TRANSITIONS = 9 # recommended
```


## 4. Predict:

```{r}
monte_carlo_runs = israeli_model$run_monte_carlo_simulation(covariates, 
                                                            to_model_state(current_state), # model uses a merged moderate/mild state
                                                            current_time = total_days_since_hospitalization,
                                                            n_random_samples = M_MONTE_CARLO_PATHS,
                                                            max_transitions = MAX_TRANSITIONS)

estimates =  data.frame(estimate = t(compute_statistics(monte_carlo_runs)))

if (COMPUTE_CONFIDENCE_INTERVALS) {
  bootstrap_sd = compute_sd_using_bootstrap(weighted_bootstrap_models,
                                            covariates, 
                                            current_state,
                                            total_days_since_hospitalization, 
                                            M_MONTE_CARLO_PATHS, 
                                            MAX_TRANSITIONS)
  
  estimates["Standard Deviation"] = as.numeric(bootstrap_sd)
  estimates["95%-lower"] = estimates$estimate - 1.96*as.numeric(bootstrap_sd)
  estimates["95%-upper"] = estimates$estimate + 1.96*as.numeric(bootstrap_sd)
}
```

### Predicted Estimates:

```{r}
formattable(estimates)
```

### CDF of remaining time in hospital:

```{r}
plot_cdf_of_remaining_time_in_hospital(monte_carlo_runs)
```

